\documentclass[a4paper]{article}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}

\newcommand{\dd}{\text{d}}
\newcommand{\ii}{\text{i}}

\newcommand{\of}[1]{\left( {#1} \right)}
\newcommand{\fracd}[2]{\frac{\dd{}{#1}}{\dd{}{#2}}}

\newcommand{\dhref}[1]{\href{#1}{#1}}
\newcommand{\mailhref}[1]{\href{mailto:#1}{#1}}

\title{Final report on ``Heterogeneity of the Suprachiasmatic
Nucleus'' {\em BO 3612/2-1}}
\author{Hanspeter Herzel  \mailhref{h.herzel@biologie.hu-berlin.de},\\
        Grigory Bordyugov \mailhref{grigory.bordyugov@gmail.com}}
\date{\today}

\begin{document}
\maketitle

\thispagestyle{empty}

\begin{abstract}
  This DFG-supported project was aiming at uncovering the rhythmic
  properties of circadian oscillator arrays of the SCN on different
  scales, ranging from single cell oscillations to ensemble effects on
  the tissue level. We report advances on several levels, including
  the conceptual (phase of entrainment), single-cell mechanistic
  (molecular workings of regulatory gene networks), and tissue-wide
  (dynamics of oscillator ensembles). The common pattern appearing
  here is that a substantial number of circadian phenomena can be well
  understood in the framework of relatively simple mathematical
  models, parameterized by few parameters such as amplitude, period
  and phase of oscillations. Additionally, the collaboration with the
  lab of Prof Toru Takumi in RIKEN resulted in discovery of Choroid
  Plexus as another strong circadian oscillator in brain with the
  rhythmicity levels comparable or even exceeding those of the SCN.
\end{abstract}

%% \tableofcontents

\newpage

\section{Overview / Zusammenfassung}
\input{overview.tex}

\newpage

\section{Main results: From modeling of single neuron oscillations to
ensembles dynamics of the SCN tissue}
Due to the space constraints for the report, below we are presenting
the most exciting (from our point of view) results obtaining during
the running time of project. A full list of publications and
manuscripts in review is given in the ``Formal'' section.

We start by discussing results on conceptual modeling of entrainment
phases in single oscillators (which can be though as an abstraction of
single cells or organisms, depending on the larger view), then present
our advances on mechanistic modelling of gene regulatory networks that
produce circadian oscillations in single cells, and look at a new
statistics that helps to characterize spatial synchronization of
rhythmic neurons on the tissue level in the SCN. We conclude by a
short report on discovering the choroid plexus as an extremely strong
circadian oscillator in the mammalian brain.


\input{phase.tex}

\input{matt.tex}

\input{mi.tex}

\input{cp.tex}

\newpage

\section{Formal}

\subsection{Published and submitted manuscripts}
A total of 7 manuscripts have been enabled by this project, five of
them having already been published and two being under review:
\cite{bordyugov2015tuning,schmal2015theoretical,kondoff2015modeling,schmal2017moran,wagner2017plant,myung2017choroid,schmal2017measuring}


\subsection{Personnel}
For the duration of the project, Dr Christoph Schmal had the position
of post-doctoral researcher at the ITB Berlin, funded by this project.

\subsection{Cooperation activities between Berlin and Tokyo}

\subsubsection{Telecom conferences}
During the three years of the project, we maintained bi-weekly Skype
conferences to synchronize our activities between Tokyo and Berlin.

\subsubsection{Visits}
The following visits between Toru Takumi's Lab at RIKEN and the
Institute for Theoretical Biology in Berlin were possible through the
project:
\begin{itemize}
  \item[-] Pia Rose visited Riken from Marth 2nd through April 25th 2015,
  \item[-] Christoph Schmal visited Riken from June 18th through July 3rd
  2014,
  \item[-] Hanspeter Herzel visited Riken from July 17th through July
  20th 2014,
  \item[-] Toru Takumi visited the ITB in Berlin on three occasions from
  2014 through 2016,
  \item With the additional support of the Einstein center
    ``Chronoi'', a three month stay of Dr. Jihwan Myung in Beriln became
    possible to continue our collaboration.
  \item[-] GB visited RIKEN from February 10th through March 1st 2014.
\end{itemize}

\subsection{Conferences}
The results of the research on the project were presented at the
Gordon Conference on Chronobiology (Newport, RI in summer 2013, also
first contact with Jihwan Myung established there), at the conference
of the Japanese Society for Mathematical Biology (Osaka, Summer 2014),
at the congress of European Biological Rhythms Society in M\"unchen
(Fall 2014), and at the meeting of SRBR (summer 2016).

\subsection{Student theses and projects}
We are reporting the following supervising activity, made possible
within this project:

Matthew Kondoff build a tool for automatically generating and
exploring regulatory networks of genes and fitted them to microarray
data of Hogenesch et al., which resulted in an R package
\cite{kondoff2015modeling}. Anna-Marie Finger and Lorena Sofia Lopez
Zepeda investigated the circadian regulation of immune system and
liver. Sungsoo Lim and Marta del Olmo conducted computer-assisted
studies of entrainment phase in several models of circadian clock.


\bibliographystyle{vancouver}
\bibliography{jg-report}
\end{document}
